Asteroid project summary

### 2022-02-04 summaries/21-11/21-11.tex
Finished spin pole clarification

Note that each sum in this equation is weighted by a factor of $a_M^\ell a_m^{\ell'}/D^{\ell+\ell'+1}$. Since $D \gg a_M \gg a_m$ in general, the leading order torques come from low $\ell$ and $\ell'$. However, $\ell'=0$ contributes nothing because the inner bracketed term is zero, and $K_{\ell' = 1} = 0$ if we choose the origin of the body-fixed axes to be the center of mass of the asteroid. The leading order of equation \ref{eqn:tidal-torque} is therefore $1/D^3$ with $\ell = 0$ and $\ell' = 2$.

@@

\subsection{Dependence on spin pole}
In addition to the physical parameters of the asteroid, the initial orientation of the asteroid spin pole strongly effects the precision of the best-fit parameters attainable from the data. For example, an asteroid with initial spin pole parallel to the orbital plane ($\beta=0$) gives Wigner-$D$ matrix elements $D^\ell_{m m'} \propto \delta_{m m'}$ and $D^\ell_{mm}={D^\ell_{-m,-m}}^*$. Note that the inner bracketed term of equation \ref{eqn:tidal-torque} transforms under $m \rightarrow -m$ as $[\cdot ] \rightarrow -(-1)^m[\cdot]^*$. Since $S_{\ell m}$ and $D_{\ell m}$ transform to their conjugates and $J_{\ell m} \rightarrow (-1)^m J_{\ell m}^*$ while none of the other terms transform, the entire equation for tidal torque transforms to the opposite of its conjugate. Since $\tau$ is invariant under this transformation, which is equivalent to a rearrangement of terms, we must have $\tau = -\tau^*$. Since $\tau \in \mathds{R}^3$, we have $\tau = 0$. Therefore the fit parameters $K_{\ell m}$ are entirely invisible when $\omega \propto \unit Z$ initially, and we expect to see decreasing fit precision near this limit.

Figure \ref{fig:spin-pole} displays the $1-\sigma$ uncertainties on the first ten fit parameters of the asymmetric reference asteroid flyby as a function of the initial direction of the spin pole. In all cases, the asteroid's initial period was set to the reference 9 hours. The data is presented mapped onto a sphere which indicates the direction of the spin pole, in the Mollweide projection. To highlight common features between the ten parameters, the average standard deviation is presented in figure \ref{fig:spin-pole-avg}. The average is weighted such that the uncertainty map for each parameter contributes an equal amount (the weight of each is set to one-tenth of its mean). Also shown in \ref{fig:spin-pole-avg} is the map of the average uncertainty with a polar projection, to show the high uncertainty near $\omega \propto \unit Z$.

@@

Figure \ref{fig:spin-pole-avg} displays other points of lack of precision: those of $\bm \omega \propto \pm \unit Y+ \frac{1}{2}\unit Z$ and to some degree $\unit X - \unit Z$. Not all of the parameters shown in figure \ref{fig:spin-pole} demonstrate this bare point, and some show lack of precision at $\bm \omega \propto -\unit Y+ \frac{1}{2}\unit Z$ as well. Better precision, however, was generally attained when aligning parallel or antiparallel to $\unit X$, the perigee vector. Initial poles with negative latitude also often relatively precise estimates. However, the average best-fit is not very volatile; it rarely grows to twice its mean over the sphere. Exactly which regions of the sphere yield better precision is likely controlled by particulars of the reference asteroid rather than fundamental properties of asteroid flybys, since no cases except when $\bm \omega \propto \unit Z$ does $\bm \tau = 0$.

@@

Figure \ref{fig:observation-gap} shows the best fit $1-\sigma$ and $2-\sigma$ uncertainty in the first order $K_{2m}$ and $\gamma_0$ and higher order $K_{3m}$ parameters as a function of the amount of perigee data $T$ that has been cut from the synthetic data set. Since torque is highest at perigee, we expect that region of the data to contain the most information about $K_{\ell m}$, and therefore that uncertainty should increase monotonically with $T$, which is observed. We also see that this increase halts after about 2 hours of data has been cut for the first order parameters, likely because the majority of the tidal torque occurs within 2 hours of perigee. The exact time of this threshold is dependent on the asteroid orbital parameters.

>>>>

\subsection{Interpretation of Tidal Torque}
\label{sec:torque-interpretation}
The rest of this paper is devoted to interpreting the properties of equation \ref{eqn:tidal-torque} through a fit to data generated with this equation of motion, but several properties are already evident through the functional form.

Firstly, each term is weighted by a factor of $a_M^\ell a_m^{\ell'}/D^{\ell+\ell'+1}$. Since $D \gg a_M \gg a_m$ in general, the leading order torques come from low $\ell$ and $\ell'$. However, $\ell'=0$ contributes nothing because the inner bracketed term is zero, and $K_{\ell' = 1} = 0$ if we choose the origin of the body-fixed axes to be the center of mass of the asteroid. The leading order of equation \ref{eqn:tidal-torque} is therefore proportional to $1/D^3$ with $\ell = 0$ and $\ell' = 2$. We therefore call $K_{2m}$ and $\gamma_0$ the ``low-order'' parameters because they are found in the leading order term of equation \ref{eqn:tidal-torque}. The next order is $K_{3m}$, and other orders after that are terms with $J_{2m}$ and $K_{2m}$ as well as $J_{00}$ and $K_{4m}$.

It is also useful to note when equation \ref{eqn:tidal-torque} is zero, since in these cases we expect to be able to determine little about the asteroid. Most important is when $\omega \parallel \unit Z$; i.e., when $\beta=0$. In this case, the Wigner-$D$ matrix elements are $D^\ell_{m m'} \propto \delta_{m m'}$ and $D^\ell_{mm}={D^\ell_{-m,-m}}^*$. Note that the inner bracketed term of equation \ref{eqn:tidal-torque} transforms under $m \rightarrow -m$ as $[\cdot ] \rightarrow -(-1)^m[\cdot]^*$. Since $S_{\ell m}$ and $D_{\ell m}$ transform to their conjugates and $J_{\ell m} \rightarrow (-1)^m J_{\ell m}^*$ while none of the other terms transform, the entire equation for tidal torque transforms to the opposite of its conjugate. Since $\tau$ is invariant under this transformation, which is equivalent to a rearrangement of terms, we must have $\tau = -\tau^*$. Since $\tau \in \mathds{R}^3$, we have $\tau = 0$.

Restricting to the first order of equation \ref{eqn:tidal-torque}, $\bm \tau$ has two other important zeros. Writing the position vector $\bm D$ in spherical coordinates $D, \theta, \phi$ and setting $\theta = \pi/2$ for simplicity, we note that $\bm \tau = 0$ when (1) $\alpha=\phi$ and $\beta = \pi/2$, and when (2) $\alpha=\phi+\pi/2$ and $\gamma$ is an integer multiple of $\pi/2$. At $\phi=0$, where $D$ is smallest and $\bm \tau$ is generally greatest, these positions correspond to (1) $\bm \omega \parallel \unit X$ and (2) $\bm \omega$ is in the $\unit Y-\unit Z$ with a condition on $\gamma$. When the latter condition on $\gamma$ is not met, $\bm \tau \propto \unit Z K_{22} \sin(2\gamma)$. Since $\gamma$ changes as the asteroid rotates and depends on the initial condition $\gamma_0$, whether the $\gamma$ condition is met at perigee is more or less random.

@@

\subsection{Dependence on spin pole}
In addition to the physical parameters of the asteroid, the initial orientation of the asteroid spin pole strongly effects the precision of the best-fit parameters attainable from the data. For example, it was pointed out that $\omega \propto \unit Z$ gives $\bm \tau = 0$ in section \ref{sec:torque-interpretation}. Since $\dot {\bm \omega} = 0$ until torque is applied to start tumbling, we expect $\bm \omega$ to be constant throughout the data set in this case, and the fit parameters $K_{\ell m}$ will be invisible. We also expect to see decreasing fit precision near this limit.

Figure \ref{fig:spin-pole} displays the $1\sigma$ uncertainties on the first ten fit parameters of the asymmetric reference asteroid flyby as a function of the initial direction of the spin pole. In all cases, the asteroid's initial period was set to the reference 9 hours. The data is presented mapped onto a sphere which indicates the direction of the spin pole, in the Mollweide projection. To highlight common features between the ten parameters, the average standard deviation is presented in figure \ref{fig:spin-pole-avg}. The average is weighted such that the uncertainty map for each parameter contributes an equal amount (the weight of each is set to one-tenth of its mean). Also shown in \ref{fig:spin-pole-avg} is the map of the average uncertainty with a polar projection, to show the high uncertainty near $\omega \propto \unit Z$.

@@

Figure \ref{fig:spin-pole-avg} confirms our expectation that the fit parameters would be uncertain near $\bm \omega \propto \unit Z$, and it displays other points of lack of precision: those of $\bm \omega \propto \pm \unit Y$. Not all of the parameters shown in figure \ref{fig:spin-pole} demonstrate this bare point: for example, $K_{20}$ and $K_{30}$ are not as affected by $\bm \omega \propto \unit Y$. The fact that these regions of imprecision line up with $\pm \unit Y$ is not guaranteed, and it is a coincidence that the reference asteroid happens to. In general, these region will be placed at other points on the equator at a location determined by the asteroid's $K_{\ell m}$ and $\gamma_0$.

This dependence on asteroid parameters is because, as section \ref{sec:torque-interpretation} points out, certain directions of $\bm \omega$ give can be shown via equation \ref{eqn:tidal-torque} to give $\bm \tau = 0$ regardless of $K_{\ell m}$. Other locations give $\bm \tau \parallel \unit Z \perp \bm \omega$, which may cause imprecision both because it restricting $\bm \tau$ to only one coordinate lowers its magnitude, and because $d/dt |\bm \omega|$ is minimized when $\bm \tau \perp \bm \omega$ and the asteroid is not tumbling. Of these low-torque locations, those that do not depend on $\gamma$ all satisfy either $\bm \omega \parallel \unit Z$ (zero torque) or $\omega_\phi = \phi + n\pi/2$, where $n$ is an integer, $\omega_\phi$ is the azimuthal coordinate of $\bm \omega$ in the global frame, and $\phi$ is the azimuthal location of the asteroid. As the asteroid encounters the central body, $\phi$ sweeps over a large range so that the low-torque loci associated with several values of $n$ affect the dynamics of the asteroid, and the degree to which the asteroid's dynamics are affected depends on $K_{ell m}$. Thus, exactly what initial spin pole direction maximizes uncertainty in the fit parameters depends on $K_{\ell m}$, but it is expected to lie in the orbital plane.

Since $K_{20}$ is not controlled by torque at this point,  and some show lack of precision at $\bm \omega \propto -\unit Y+ \frac{1}{2}\unit Z$ as wel. Better precision, however, was generally attained when aligning parallel or antiparallel to $\unit X$, the perigee vector. Initial poles with negative latitude also often relatively precise estimates. However, the average best-fit is not very volatile; it rarely grows to twice its mean over the sphere. Exactly which regions of the sphere yield better precision is likely controlled by particulars of the reference asteroid rather than fundamental properties of asteroid flybys, since no cases except when $\bm \omega \propto \unit Z$ does $\bm \tau = 0$.

@@

Figure \ref{fig:observation-gap} shows the best fit $1\sigma$ and $2\sigma$ uncertainty in the first order $K_{2m}$ and $\gamma_0$ and higher order $K_{3m}$ parameters as a function of the amount of perigee data $T$ that has been cut from the synthetic data set. Since torque is highest at perigee, we expect that region of the data to contain the most information about $K_{\ell m}$, and therefore that uncertainty should increase monotonically with $T$, which is observed. We also see that this increase halts after about 2 hours of data has been cut for the first order parameters, likely because the majority of the tidal torque occurs within 2 hours of perigee. The exact time of this threshold is dependent on the asteroid orbital parameters.
